### 2024-10-22 paper/paper.qmd
Basic steps of the model

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - David Qi
  - Haobo Ren
  - Xinrui Xie
thanks: "Code and data are available at: https://github.com/HaoboRrrr/USA_Election"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

# Data {#sec-data}

## Overview
Josh Pasek have discussed how to the use of Aggregation and Predictive Modeling on polling can increase prediction effectiveness[@Pasek_2015]. In this article, we will make use of predictive modeling to create a model that aim to predict the support rate of Kamala Harris in the 2024 United States presidential election using a linear regression model trained form past polling results and Dow Jones Industrial Average.

## Measurement

## Outcome variables

## Predictor variables

# Model

We will use a linear model to predict the general polling result of Harris.

## Model set-up
For each poll, we assumed they are being sampled similarly. \
Let the support rate for Harris at day $t$ be denoted by $R_t$. And we denote the Dow Jones Industrial Average at day $t$ by $D_t$.
Then we will use linear regression model: 
$$
R_t = D_{t-60} +  D_{t-90} + \epsilon_t
$$
Where $\epsilon_t$ is an independent, normally distributed error term.

## Model justification

We expect a positive relationship between the support rate of Harris and Dow Jones Industrial Average 60 and 90 days before. Since people tend to support the party in power when the economy is better, where the condiction of the economy is reflected by $D_t$.

TODO: find paper to support this.
finishing reading https://dspacemainprd01.lib.uwaterloo.ca/server/api/core/bitstreams/4dfe49d7-72b7-44e5-8fdc-8a5cf0b77b24/content

# Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

\newpage

# Appendix 1: Polling methodology

# Appendix 2: Ideal methodology and survey

# References