### 2024-11-04 paper/paper.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - David Qi
  - Haobo Ren
  - Xinrui Xie
thanks: "Code and data are available at: https://github.com/HaoboRrrr/USA_Election"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

# Data {#sec-data}

## Overview
Josh Pasek have discussed how to the use of Aggregation and Predictive Modeling on polling can increase prediction effectiveness[@Pasek_2015]. In this article, we will make use of predictive modeling to create a model that aim to predict the support rate of Kamala Harris in the 2024 United States presidential election using a linear regression model trained form past polling results and Dow Jones Industrial Average.

## Measurement

## Outcome variables

## Predictor variables

# Model

We will use a linear model to predict the general polling result of Harris.

## Model set-up
For each poll, we assumed they are being sampled similarly. \
Let the support rate for Harris at day $t$ be denoted by $R_t$. And we denote the Dow Jones Industrial Average at day $t$ by $D_t$.
Then we will use linear regression model: 
$$
R_t = D_{t-60} +  D_{t-90} + \epsilon_t
$$
Where $\epsilon_t$ is an independent, normally distributed error term.

## Model justification

We expect a positive relationship between the support rate of Harris and Dow Jones Industrial Average 60 and 90 days before. Since people tend to support the party in power when the economy is better, where the condiction of the economy is reflected by $D_t$.

TODO: find paper to support this.
finishing reading https://dspacemainprd01.lib.uwaterloo.ca/server/api/core/bitstreams/4dfe49d7-72b7-44e5-8fdc-8a5cf0b77b24/content

# Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

\newpage

# Appendix 1: Polling methodology

# Appendix 2: Ideal methodology and survey

## Objective
The survey aims to capture voter intention and mood across important demographic and geographic divisions in order to predict the result of the US presidential election. This survey's methodology employs stratify sampling, effective respondent recruitment strategy and data validation techniques to ensure the accuracy of prediction.

## Sampling Approach
This survey employs Stratified Random Sampling, which involves divide the target population(In this case, the people who are eligible to vote in US) into subgroups(strata) based on their demographics characteristic. To simulate the national electorate, stratify the sample according to age, gender, race/ethnicity, education level, urban/rural domicile, and region. Also refer to the US Census data to determine the proportion of each strata, for example: if the census data shows there are 20% US citizen are between the age of 20 to 30, then there should be 20% of respondent in this age range. We aim for a starting sample of five thousand people since there were budget constraint.

## Respondent Recruitment
Partner with an online panel provider to find respondents that match each stratum’s specifications.Specify quotas to ensure that, for instance, you wish to have 18% of respondents from the Midwest or 25% aging between 30 - 40. Then randomly select respondent within each strata. To promote greater response rates and retention, provide a small monetary reward, such as $5 for each completed survey. To improve inclusivity and take into consideration differences in internet availability, use both phone and online surveys.

## Data Validation
* De-duplication: To keep distinct replies between waves, make sure respondents don't participate more than once each wave.

* Quality checks: highlight straight-lining, inconsistent replies, and other low-effort answers.

## Budget Expenditure
* \$50k - \$60k for panel provider
* \$25k for incentives
* \$10k for Telephone survey services
* \$5k for data analysis

## Survey Structure
This survey will focus only on collecting participants' preferences for presidential candidates, not senate elections.Therefore, this survey only asks participants' choice of their favored presidential candidates. This design can make the entire questionnaire more simplify, while also collecting key data for predicting the presidential election.

Screener Section: Screening out people who are eligible to vote in each state. This can filter out participants who are not eligible to vote, thereby improving the accuracy of the prediction results.

* Are you currently registered to vote?
  - Yes
  - No

* Which state are you currently registered to vote in?
  - (Drop down box)

Presidential Ballot Section: Ask participants whether they would vote, who they would vote for, and how confident they are about that. This section can be used to find out the proportion of votes each candidate received among the participants.

* Do you plan to vote in the upcoming Presidential election?
  - Yes, I plan to vote in person on election day
  - Yes, I plan to vote in person early
  - Yes, I plan to vote by mail
  - No, I do not plan to vote

* If the election were held today, who would you vote for?
  - Democrat Kamala Harris
  - Republican Donald Trump
  - A third party / Independent candidate
  - Unsure

* How certain are you about your choice
  - Scale 1 to 5

Demographics Section: This section collects detailed demographic characteristics of the participants to accurately predict the nation-wide election outcome.

* Your gender identification
  - Female
  - Male
  - Other

* Age Range
  - 18 to 24
  - 25 to 29
  - 30 to 39
  - 40 to 49
  - 50 to 59
  - 60 or older

* Which of the following best describes your race or ethnicity
  - White or Caucasian
  - Black or African American
  - Hispanic
  - Asian or Pacific Islander
  - Native American
  - Other

* How would you describe your educational level
  - High School or less
  - College Graduate
  - Post-graduate degree
  - Unsure

* Employment Status
  - Full time
  - Part time
  - Unemployed

* What's your estimated household income?
  - Under $30,000
  - \$30k - \$49,999
  - \$50k - \$99,999
  - \$100k - \$199,999
  - \$200k - \$249,999
  - \$250k or more
  - Unsure

* Do you consider yourself to be lesbian, gay, bisexual, transgender, queer or
questioning?
  - Yes
  - No
  - Unsure
  - Prefer not to answer

* Would you best describe the neighborhood or area you live in as
  - Urban
  - Suburban
  - Rural
  - Unsure

* What is your religious background
  - Religious
  - Spiritual
  - Not religious or spiritual

Link to the survey: https://forms.gle/spJjASENRQ6nKK1E6

# References